---
title: "Project1"
author: "Georges Ip"
date: "2/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries
```{r message=FALSE}
library(tidyverse)
library(tseries) 
library(car)
library(forecast) 
require(stats)
library(vars)
library(MASS) 
require(stats4) 
require("datasets") 
require(graphics) 
library(RColorBrewer) 
library(dynlm)
library(ggcorrplot)
library(olsrr)
library(readxl)
library(ISLR)
library(dplyr)
library(tidyverse)
library(class)
library(gridExtra)
library(caTools)
library(boot)
library(caret)
library(e1071)
```


Import datasets
```{r}
train <- read_csv("train.csv")
test <- read_csv("test.csv")
```

Explore dataset
```{r}
attach(train)
head(train)
colSums(is.na(train))
```



There are quite a few missing observations in Age and Cabin, as well as 2 observations in Embarked. Some of the indicator variables are also classified as num. 

```{r}
#First, deal with the factor variables
factors <- c("Survived", "Pclass", "Sex", "Embarked")
train[,factors] <- lapply(train[,factors], factor)
```



Now we have convertd to factor variables, we can conduct some preliminary descriptive analysis. A correlation plot will be more difficult to interpret as we have a few factor variables. 
```{r}
#First look at Passenger Class
table1 <- table(Pclass,Survived)
ftable(round(prop.table(table1), 3))
```



From this table, we can see that there is quite a clear relationship between survival and Pclass, with a higher proportion of survivors in better passenger classes. 
```{r}
#Now look at Sex
table2 <- table(Sex, Survived)
ftable(table2)
```



It is evident that a higher proportion of females survived compared to males

```{r}
#Look at the combined table
table3 <- table(Pclass, Sex, Survived)
ftable(table3)
```


This table is slightly more diifficult to interpret, but ones notable takeaway is that both trends we identified within Pclass and Sex still hold. This could be variables used in classification. We now turn our attention to numerical variables.



```{r}
#Add a histogram to see if ages affects survival
hist(Age[Survived == 0], density = 50, col = 2, angle = 0)
hist(Age[Survived == 1], add=TRUE, density = 25, col = 3, angle = 0)
```


```{r}
#Subset the train set
set.seed(1)
train.vector = sample(1:891,500)
train.set = train[train.vector,]
test.set = train[-train.vector,]
#validation
validation.set = train$Survived[-train.vector]
```


```{r}
#Clean test set as well and add a Survived column 
factors <- c("Pclass", "Sex", "Embarked")
test[,factors] <- lapply(test[,factors], factor)
```



```{r}
#Try LDA
lda.fit=lda(Survived~Sex + Age + Pclass + SibSp + Parch + Fare, data=train, na.action = na.omit)
lda.pred=predict(lda.fit, test.set)
lda.class=lda.pred$class
table(lda.class ,validation.set)
mean(lda.class==validation.set)
#We can also apply a 50% threshold to the data
sum(lda.pred$posterior[,1]>=.5)
sum(lda.pred$posterior[,1]<.5)
```



```{r}
mean(train$Survived)
```


```{r}
train$Sex01 <- train$Sex == "female"
test$Sex01 <- test$Sex == "female"
```





```{r}
glm1 <- glm(Survived ~ Pclass + Sex01 + Age + SibSp + Parch + Fare, data = train, family = "binomial")
glm1
```

```{r}
influenceIndexPlot(glm1, vars="Cook")
```

```{r}
resettest(glm1, power=2)
```

```{r}
glm2 <- glm(Survived ~ Pclass + Sex01 + Age + I(Age^2) + SibSp + Parch + Fare, data = train, family = "binomial")
glm2
```


```{r}
resettest(glm2, power = 2)
```

```{r}
library(leaps)
```



```{r}
ss <- regsubsets(Survived ~ Pclass + Sex01 + Age + I(Age^2) + SibSp + Parch + Fare, data = train, method = "exhaustive", nbest = 2)
subsets(ss,statistic="cp",legend=F, ylim = c(0,20))

```
```{r}
glm3 <- glm(Survived ~ Pclass + Sex01 + Age + I(Age^2) + SibSp, data = train, family = "binomial")
glm3
```

```{r}
AIC(glm1,glm2,glm3)
```





```{r warning=FALSE}
lda1 <- lda(Survived ~ Pclass + Sex01 + Age + SibSp + Parch + Fare, data = train, na.action = na.exclude)

lda1pred <- predict(lda1, train, type = "response", na.action = na.exclude)

mean(lda1pred$class == train$Survived, na.rm = T)
```

```{r warning=FALSE}
lda2 <- lda(Survived ~ Pclass + Sex01 + Age + I(Age^2) + SibSp, data = train, na.action = na.exclude)

lda2pred <- predict(lda2, train, type = "response", na.action = na.exclude)

mean(lda2pred$class == train$Survived, na.rm = T)
```


```{r warning=FALSE}
qda1 <- qda(Survived ~ Pclass + Sex01 + Age + SibSp + Parch + Fare, data = train, na.action = na.exclude)

qda1pred <- predict(qda1, train, type = "response", na.action = na.exclude)

mean(qda1pred$class == train$Survived, na.rm = T)
```

```{r warning=FALSE}
qda2 <- qda(Survived ~ Pclass + Sex01 + Age + I(Age^2) + SibSp , data = train, na.action = na.exclude)

qda2pred <- predict(qda2, train, type = "response", na.action = na.exclude)

mean(qda2pred$class == train$Survived, na.rm = T)
```

```{r}
library(lava)
```


```{r}
cv.err <- cv.glm(train, glm3)
cv.err$delta
```

